---
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```
# tc.sensors  

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing) ![check-package](https://github.com/Metropolitan-Council/tc.sensors/workflows/check-package/badge.svg)
[![Website MnDOT JSON](https://img.shields.io/website-up-down-green-red/http/shields.io.svg)](http://data.dot.state.mn.us:8080/trafdat/metro/2018/20181021/5474.c30.json)
<!-- badges: end -->
  
## Overview
A package for pulling data for Minnesota Department of Transportation (MnDOT) loop detectors installed on the Minnesota Freeway system in 30-second interval measurements of occupancy and volume, data which are pushed daily to a public JSON feed.  

## Installation  

```{r}
remotes::install_github("Metropolitan-Council/tc.sensors")
```

## Documentation

To access documentation and for help on how to use the package, run `?<FUNCTION-NAME>` (e.g. `?pull_sensor`, `?pull_configuration`, `?pull_sensor_ids`).  Access vignettes in the vignettes file to see examples of end-to-end workflows for pulling and storing data locally en masse.  Check back for a vignettes for calculating speeds, reference speeds, delay, and VMT from the resulting files.

## Contributors  


* **Maintainer** Liz Roten (liz.roten@metc.state.mn.us)  
* Nicole Sullivan (nicole.sullivan@metc.state.mn.us)


## Code of Conduct
  
  Please note that the tc.sensors project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
