---
title: "Calculate speed and delay"
author: Liz Roten
output: rmarkdown::html_vignette
bibliography: references.bib
csl: apa-single-spaced.csl
nocite: '@*'
vignette: >
  %\VignetteIndexEntry{Calculate Speed and Delay}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,
  warning = FALSE,
  message = FALSE
)

options(scipen = 9999)
```

## Pull sensor data

First, we will pull the data for three random sensors from March 3 to March 16, 2019. 

```{r}
library(tc.sensors)
library(tictoc)
library(dplyr)
library(ggplot2)
library(purrr)
library(furrr)
library(cowplot)
library(stringr)
```


```{r, cache=TRUE}
sensor_config <- pull_configuration() %>%
  dplyr::filter(
    corridor_route == "T.H.610", # single out one corridor
    r_node_n_type == "Station", # only station node types
    detector_category == "" # only mainline detectors
  )

date_range <- seq(as.Date("2019/03/03"), as.Date("2019/03/16"), by = "days")

plan(multiprocess)

tic() # Start the timer
sensor_data <- future_map(sensor_config$detector_name,
  .f = function(det_name) {
    future_map_dfr(seq(1:length(date_range)), .f = function(x) {
      tc.sensors::pull_sensor(det_name, date_range[[x]])
    })
  }
)
toc()
```


Clean data

```{r}
clean_sensor_data <- purrr::map(sensor_data,
  scrub_sensor,
  interval_length = NA
) # replace
```

## Aggregate sensor data and calculate speed  

To calculate speed and evaluate our speed calculation, we are going to look at two different aggregation intervals:  10 minutes and 60 minutes (or 1 hour). 

To avoid getting unreasonable speed values, we will set an occupancy percentage threshold at 0.002, or 0.2%. This functions somewhat like a sample size cut-off; if the occupancy is too low, its more likely that the speed value returned will be invalid. 

```{r}
agg_ten <- purrr::map(clean_sensor_data,
  aggregate_sensor_data,
  config = sensor_config,
  interval_length = 10 / 60,
  occupancy_pct_threshold = 0.002
) %>%
  purrr::map_dfr(add_day_type)

agg_hour <- purrr::map(clean_sensor_data,
  aggregate_sensor_data,
  config = sensor_config,
  interval_length = 1,
  occupancy_pct_threshold = 0.002
) %>%
  purrr::map_dfr(add_day_type)
```

The NA values here aren't evenly distributed across all starting hours and day types. There is a greater percentage of null values in the early morning on both weekdays and weekends, which aligns with our knowledge that there is relatively low traffic during these hours. 

```{r}
# > 10min speed summary
summary(agg_ten$speed)
# > 10 minute speed NA percentage
summary(agg_ten$speed)[7] / nrow(agg_ten)

# > 1 hour speed summary
summary(agg_hour$speed)
# > 1 hour minute speed NA percentage
summary(agg_hour$speed)[7] / nrow(agg_hour)
```

We also see this in the 

```{r}
cowplot::plot_grid(
  agg_ten %>%
    group_by(day_type, hour) %>%
    summarise(
      pc_n = sum(is.na(speed)) / length(speed),
      .groups = "keep"
    ) %>%
    ggplot() +
    geom_col(
      mapping = aes(
        x = hour,
        y = pc_n,
        fill = day_type
      ),
      alpha = 0.7,
      position = position_dodge(width = 0.3),
    ) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_viridis_d("Day Type", option = "magma", begin = 0.2, end = 0.7) +
    theme_minimal() +
    labs(
      y = "% null speed observations",
      x = "Start hour",
      title = "10 minute aggregation"
    ) +
    theme(legend.position = "bottom"),
  agg_hour %>%
    group_by(day_type, interval_bin) %>%
    summarise(
      pc_n = sum(is.na(speed)) / length(speed),
      .groups = "keep"
    ) %>%
    ggplot() +
    geom_col(
      mapping = aes(
        x = interval_bin,
        y = pc_n,
        fill = day_type
      ),
      position = position_dodge(width = 0.3),
      alpha = 0.7
    ) +
    scale_fill_viridis_d("Day Type", option = "magma", begin = 0.2, end = 0.7) +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(limits = range(agg_ten$hour)) +
    theme_minimal() +
    labs(
      y = "",
      x = "Start hour",
      title = "1 hour aggregation"
    ) +
    theme(legend.position = "bottom")
)
```

Next, we will plot the relationship between speed, volume, and occupancy for both aggregation levels. The plots below show that the larger interval size seems to stabilize speed values. 

```{r}
plot_ten <- ggplot() +
  geom_point(
    data = agg_ten %>%
      arrange(speed),
    aes(
      y = occupancy.pct,
      x = speed,
      fill = scale(volume.sum),
      color = scale(volume.sum)
    ),
    alpha = 0.5,
    size = 3
  ) +
  scale_fill_viridis_c(stringr::str_wrap("Relative volume", 5), option = "magma", aesthetics = c("fill", "color")) +
  scale_y_continuous("Occupancy %",
    labels = scales::percent,
    limits = range(agg_ten$occupancy.pct)
  ) +
  scale_x_continuous("Speed (mph)", limits = range(agg_ten$speed)) +
  theme_minimal() +
  labs(title = "10 minute aggregation") +
  theme(legend.position = "bottom")



plot_hour <- ggplot() +
  geom_point(
    data = agg_hour %>%
      arrange(speed),
    aes(
      y = occupancy.pct,
      x = speed,
      fill = scale(volume.sum),
      color = scale(volume.sum)
    ),
    alpha = 0.5,
    size = 3
  ) +
  scale_fill_viridis_c(stringr::str_wrap("Relative volume", 5), option = "magma", aesthetics = c("fill", "color")) +
  scale_y_continuous("",
    labels = scales::percent,
    limits = range(agg_ten$occupancy.pct)
  ) +
  scale_x_continuous("Speed (mph)", limits = range(agg_ten$speed)) +
  theme_minimal() +
  labs(title = "1 hour aggregation") +
  theme(legend.position = "bottom")


cowplot::plot_grid(plot_ten, plot_hour)
```

## Delay  

*Delay* is the difference in the time it takes to traverse a given distance at the reference speed and the comparison speed. Reference speed is typically the speed during free-flow periods. Delay is highly dependent on the reference speed.  

>The use of free-flow speeds as a benchmark is understandable since it can be seen as a replicable and readily useable “objective” figure. Problems arise, however, when the difference between free flow speeds and experienced speeds are labelled “delay”. Delay is the technically correct term for the difference but is often semantically misconstrued as referring to an attainable target for peak hour travel – e.g. “zero-delay”. It should be noted that, most dynamic cities cannot afford to deliver free-flow speeds at peak hours, nor would they want to live with a road network that could deliver these speeds at peak hours. Discourse based on delay as measured in reference to free-flow travel times can thus be biased towards an unattainable and likely undesirable congestion management goal (e.g. zero delay). In this case, the use of an outcome neutral indicators or indices is preferable.  - *@world_bank_cairo_2013, Annex 9*

To calculate the amount of time (in hours) it takes to traverse a given distance, we need to know the distance length (in miles) and the speed (in miles per hour). 

$Time = \frac{Distance}{Speed}$

We can safely assume that the distance between sensors is 0.5 miles. 


## Relevant reports and references  

<div id="refs"></div>

<footer>
*Last updated by Liz Roten, `r Sys.Date()`*
</footer>
