---
title: "Compare RTMC Congestion Report with loop detector data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{compare_congestion_metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  # fig.width = 8,
  # fig.height = 4,
  # dev = "CairoPNG",
  fig.path = "../man/figures/compare_congestion_metrics-"
)

options(scipen = 9999)

th <- ggplot2::theme_minimal() +
  councilR::council_theme(use_showtext = TRUE,
                          size_header = 16,
                          size_axis_title = 12) +
  ggplot2::theme(legend.position = "bottom")

```

```{r setup}
library(tc.sensors)
library(sf)
library(data.table)
library(dplyr)
library(tidyr)
library(janitor)
library(leaflet)
library(htmltools)
library(reactable)
```

Our goal is to link the freeway congestion report shapefile with our stations.  

## Regional Traffic Management Center (RTMC) 2019 Congestion report data  

The freeway congestion data represents the average of non-holiday Tuesdays, Wednesdays, and Thursdays in October 2019.  

- `am_sbc` indicates the number of five-minute increments where average speeds fell below 45mph from 5am to 10am  
- `pm_sbc` indicates the number of five minute increments where average speeds fell below 45mph between 2pm to 7pm
- `am_pm_full` sum of AM and PM

There are (60 minutes) * 5 hours = 300 minutes minutes in each time period. This means there are 300 min./5 min. = 60 five-minute increments in each 5-hour period. 

- AM = 60 five-minute increments 
- PM = 60 five-minute increments  
- Daytime (5am-7pm) = 168 five-minute increments

```{r}
rtmc_congest <- read_sf("data_cache/2019_freeway_congestion/2019_freeway_congestion.shp") %>%   clean_names() %>% 
  sf::st_set_crs(26915) %>% 
  st_transform(4326) %>% 
  mutate(hover_text = stringr::str_wrap(paste0(
                            roadway, " ", direction1, 
                             " | Station: ", sta_1), 3)) %>% 
  mutate(am_perc = round(am_sbc/60, 4),
         pm_perc = round(pm_sbc/60, 4),
         am_pm_perc = round(am_pm_full/120, 4)) %>% 
    select(station = sta_1,
    roadway, direction = direction1,
         am_perc, pm_perc, am_pm_perc, 
    hover_text)

  
  

reactable(rtmc_congest, highlight = TRUE, bordered = TRUE,
          compact = TRUE)
```

```{r}
congest_pal <- colorNumeric(palette = "viridis", domain = rtmc_congest$am_pm_full)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = rtmc_congest,
              fillColor = ~congest_pal(am_pm_full),
              fillOpacity = 0.8, 
              smoothFactor = 0.2,
              stroke = FALSE,
              label = ~(hover_text),
              labelOptions = labelOptions(textsize = 30))
```

## Link stations with sensors  

```{r}
sensor_config <- pull_configuration() %>% 
  filter(detector_abandoned == "f",
         r_node_n_type == "Station",
         detector_category == "",
         r_node_staton_id %in% rtmc_congest$station) 
```

```{r}
# sensor_data <- pull_sensor_range("2019-10-01",
# "2019-10-31", sensor_config)
# saveRDS(sensor_data, "data_cache/october_2019_stations.RDS")
sensor_data <- readRDS("data_cache/october_2019_stations.RDS")
```


```{r}
# agg_five <- furrr::future_map(sensor_data,
#                                       aggregate_sensor,
#                                       config = sensor_config,
#                                       interval_length = 5 / 60,
#                                       occupancy_pct_threshold = 0.002,
#                                       interpolate_missing = TRUE
# )  %>%
#   furrr::future_map(add_day_type)
# saveRDS(agg_five, "data_cache/october_2019_agg_five.RDS")
# # 
# agg_five_midweek <- agg_five %>%
#   furrr::future_map_dfr(.f = function(x){
#     x[day_of_week %in% c("Tuesday", "Wednesday", "Thursday")][
#       hour %in% c(seq(5,10), seq(14,19)),][
#         , period := ifelse(hour > 10, "PM", "AM")]
#   })
# 
# saveRDS(agg_five_midweek, "data_cache/october_2019_agg_five_midweek.RDS")

agg_five <- readRDS("data_cache/october_2019_agg_five.RDS")
agg_five_midweek <- readRDS("data_cache/october_2019_agg_five_midweek.RDS")
```



```{r}
length(unique(sensor_config$r_node_station_id))

sensor_rtmc <- right_join(rtmc_congest,
                          sensor_config, by = c("station" = "r_node_station_id",
                                              "direction" = "corridor_dir" ))

```

