---
title: "Key Congestion Metrics"
author: Liz Roten, `r Sys.Date()`
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
csl: apa-single-spaced.csl
nocite: '@*'
vignette: >
  %\VignetteIndexEntry{Key Congestion Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
eval = FALSE,
warning = FALSE,
message = FALSE,
cache = TRUE,
fig.width = 8,
fig.height = 4,
fig.path = "../man/figures/key_congestion_metrics-"
)

options(scipen = 9999)
```


```{r setup}
library(tc.sensors)
library(purrr)
library(furrr)
library(tictoc)
library(data.table)
library(dplyr)
library(ggplot2)
library(tidyr)
library(leaflet)
```

## Data  

```{r}
sensor_config <- pull_configuration()[
  detector_category == "",][
    r_node_n_type == "Station",][
      corridor_route == "I-94",]
# only include mainline, station sensors on I-94

```

```{r}
date_range <- seq(as.Date("2018/10/02"), as.Date("2018/10/09"), by = "days")
plan(multiprocess)

tic() # Start the timer
sensor_data <- future_map(sensor_config$detector_name,
                          .f = function(det_name) {
                            future_map_dfr(seq(1:length(date_range)), .f = function(x) {
                              pull_sensor(det_name, date_range[[x]])
                            })
                          }
)
toc()
```

## Metrics  

There are a few key metrics we want to evaluate when deciding how to define congestion. 

### Speed   

Currently, the RTMC defines congestion as an operating speed under 45 miles per hour (@minnesota_department_of_transportation_mndot_2019). If we change this threshold to a different speed, we can dramatically affect how much of a given road is congested.  

There are a 

```{r}

```


We will select only observations on midweek days in the study hours. 

```{r}
agg_five <- furrr::future_map(sensor_data,
                              aggregate_sensor,
                              config = sensor_config,
                              interval_length = 5 / 60,
                              occupancy_pct_threshold = 0.002,
                              interpolate_missing = TRUE
)  %>% 
  furrr::future_map(add_day_type) %>% 
  furrr::future_map_dfr(.f = function(x){
    x[day_of_week %in% c("Tuesday", "Wednesday", "Thursday")][
      hour %in% c(seq(5,10), seq(14,19)),][
        , period := ifelse(hour > 10, "PM", "AM")]
  })


agg_five_clean <- agg_five[
      , congested.calc := ifelse(speed.calc <= 45, TRUE, FALSE)][
        , congested.prov := ifelse(speed.median <= 45, TRUE, FALSE)][
          , congested.any := ifelse(congested.calc != TRUE | congested.prov == TRUE,
                                    TRUE, FALSE)
        ]
  
```


```{r}
agg_hour <- furrr::future_map(sensor_data,
                              aggregate_sensor,
                              config = sensor_config,
                              interval_length = 1,
                              occupancy_pct_threshold = 0.003,
                              interpolate_missing = TRUE
)  %>% 
  furrr::future_map(add_day_type)  %>% 
  furrr::future_map_dfr(.f = function(x){
    x[day_of_week %in% c("Tuesday", "Wednesday", "Thursday")]
  })
```

```{r}
agg_five_clean %>% 
  left_join(sensor_config, by = c("sensor" = "detector_name")) %>% 
  group_by(r_node_station_id) %>% 
  count(r_node_station_id, congested.prov) %>% 
  spread(congested.prov, n, ) %>% 
  mutate(percent_congested = round(`TRUE`/sum(`FALSE`, `TRUE`, `<NA>`, na.rm = T) * 100, 2))
```


```{r}
ggplot() +
  geom_jitter(agg_hour,
              mapping = aes(x = speed.median,
                            y = speed.calc)) +
  geom_abline(slope = 1) +
  theme_minimal()

```

```{r}
sensor_config_length <- add_distance(sensor_config)
setkey(sensor_config_length, detector_name)

sensor_config_length[, .(`sum(distance)` = sum(distance))
                     # keyby = .(corridor_route, corridor_dir)
    ]
```


```{r}
setkey(agg_five_median_speed, sensor) 

summary(agg_five_median_speed$median_speed)

agg_five_congested_stations <- agg_five_median_speed[,congested_status := ifelse(median_speed <= 45, TRUE, FALSE)]

summary(agg_five_congested_stations$congested_status)

agg_five_congested_stations[sensor_config_length][,node_congested_status := ifelse(congested_status == TRUE, TRUE, FALSE), keyby = .(r_node_station_id, period, date)][,.(r_node_station_id, period, date, distance, median_speed, node_congested_status)] %>% unique() %>% View
```

```{r}
summary(agg_five$congested_status)
```

