---
title: "Link with NPMRDS"
output: rmarkdown::html_vignette
author: Liz Roten, `r Sys.Date()`
bibliography: ../inst/REFERENCES.bib
csl: apa-single-spaced.csl
nocite: '@*'
vignette: >
  %\VignetteIndexEntry{Link with NPMRDS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  fig.width = 7,
  fig.height = 4,
  fig.path = "../man/figures/link_with_npmrds-"
)

options(scipen = 9999)
```

```{r setup}
library(tc.sensors)
library(rritis)
library(sf)
library(tictoc)
library(dplyr)
library(ggplot2)
library(furrr)
library(purrr)
library(data.table)
```


## Read in NPMRDS data  

The data we are using is from March 2019 for passenger and commercial vehicles.  

```{r}
tictoc::tic()
passenger_march <- import_npmrds(npmrds_data_location = "../../data_ritis/data_npmrds/massive_data_downloader/MPO-March2019-Passenger-Commerical-seconds/MPO-March2019-Passenger-Commerical-seconds.csv")[
  , `:=`(
    start_hour = data.table::hour(measurement_tstamp),
    start_date = as.Date(measurement_tstamp)
  )][
    , `:=`(study_period = data.table::fcase(
      start_hour %in% seq(5, 10), "Morning rush",
      start_hour %in% seq(14, 19), "Evening rush",
      default = "None"
    ))
  ]
tictoc::toc()
```


```{r}
inrix_spatial <- import_npmrds_inrix_spatial(
  state = "MN",
  year = "2019",
  location = "../../data_ritis/data_npmrds/inrix_shapefiles/"
) %>%
  dplyr::filter(tmc %in% unique(passenger_march$tmc_code)) # filter to only include metro roads
```


## Prepare loop sensor data for comparison  

Spatial join sensors with INRIX spatial data
```{r}
sensor_config <- pull_configuration() %>% 
  filter(detector_abandoned == "f") %>% 
  mutate(lat = as.numeric(r_node_lat),
         lon = as.numeric(r_node_lon)) %>% 
  st_as_sf(
  coords = c("lon", "lat")
) %>% 
  st_set_crs(4326)


inrix_nearest_join <- st_join(inrix_spatial, sensor_config, join = st_nearest_feature)


inrix_config_intersect <- st_intersection(sensor_config,
                                          inrix_spatial, tolerance = 2)

inrix_sensor_ids <- inrix_config_intersect %>% 
  select(detector_name, tmc)
```

```{r}
ggplot() +
  geom_sf(data = inrix_spatial) +
  geom_sf(data = inrix_config_intersect,
          color = "blue") +
  theme_minimal()
```

## Pull and process loop sensor data  

```{r}
date_range <- seq(as.Date("2019/03/01"), as.Date("2019/03/31"), by = "days")

plan(multiprocess)

tic() # Start the timer
sensor_data <- future_map(inrix_sensor_ids$detector_name,
  .f = function(det_name) {
    future_map_dfr(seq(1:length(date_range)), .f = function(x) {
      tc.sensors::pull_sensor(det_name, date_range[[x]])
    })
  }
)
toc()
```

Aggregate sensor data to 5 minute intervals.  

```{r}
plan(multiprocess)
tictoc::tic()

clean_sensor_data <- furrr::future_map(sensor_data,
  scrub_sensor,
  interval_length = NA
)

agg_five <- furrr::future_map(clean_sensor_data,
  aggregate_sensor,
  config = sensor_config,
  interval_length = 5 / 60,
  occupancy_pct_threshold = 0.002,
  interpolate_missing = TRUE
) %>%
  furrr::future_map_dfr(add_day_type)
tictoc::toc()
```

Join with configuration. 

```{r}
sample_date <- "2019-03-12"

passenger_march_match <- passenger_march[
  tmc_code %in% inrix_sensor_ids$tmc,][
    start_date == sample_date,][
      ,measurement_tstamp := as.character(measurement_tstamp)][
        ,.(tmc_code, speed, measurement_tstamp, data_density)]

agg_five_dt <- as.data.table(agg_five, keys = c("detector_name"))[
  date == sample_date,][
    , start_datetime := as.character(start_datetime)][
      , .(sensor, start_datetime, volume.sum, speed)]

massive_join <- left_join(agg_five_dt, inrix_sensor_ids, by = c("sensor" = "detector_name"))  %>% 
  left_join(passenger_march_match, suffix = c(".loop", ".inrix"), by = c("tmc" = "tmc_code",
                                          "start_datetime" = "measurement_tstamp")) %>% 
  mutate(start_datetime = as.POSIXct(start_datetime),
         late_night = hour(start_datetime) %in% c(22:23, 0:4))
```

```{r}
ggplot() +
  geom_jitter(data = massive_join,
             aes(x = speed.loop,
                 y = speed.inrix,
                 color = volume.sum),
             alpha = 0.5) +
  geom_smooth(data = massive_join,
              aes(x = speed.loop,
                  y = speed.inrix)) +
  scale_color_viridis_c(option = "magma") +
  theme_minimal() +
  labs(x = "Loop detector speed (mph)",
       y = "INRIX speed (mph)",
       color = "Volume",
       title = "INRIX and loop detector speeds")
```

To get a better picture of data periods with low volumes, we will mark late night observations as those starting between midnight and 4am and 10pm to 11pm.

```{r}
ggplot() +
  geom_jitter(data = massive_join,
             aes(x = speed.loop,
                 y = speed.inrix,
                 color = volume.sum),
             alpha = 0.5) +
  geom_smooth(data = massive_join,
              aes(x = speed.loop,
                  y = speed.inrix)) +
  facet_wrap(~late_night) +
  scale_color_viridis_c(option = "magma") +
  theme_minimal() +
    labs(x = "Loop detector speed (mph)",
       y = "INRIX speed (mph)",
       color = "Volume",
       title = "INRIX and loop detector speeds by late night")

```

## MnDOT AADT count locations and segments

```{r}
## point locations
temp <- tempfile()
download.file("https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_dot/trans_aadt_traffic_count_locs/gpkg_trans_aadt_traffic_count_locs.zip", temp)

locations <- read_sf(unzip(temp, "trans_aadt_traffic_count_locs.gpkg")) %>% 
  st_transform(4326) 
fs::file_delete("trans_aadt_traffic_count_locs.gpkg")
```

```{r}

temp <- tempfile()
download.file("https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_dot/trans_aadt_traffic_segments/gpkg_trans_aadt_traffic_segments.zip", temp)

segments <- read_sf(unzip(temp, "trans_aadt_traffic_segments.gpkg")) %>% 
  st_transform(4326)

fs::file_delete("trans_aadt_traffic_segments.gpkg")
fs::file_delete(temp)

counti <- tigris::counties(state = 27, year = 2018, 
                           class = "sf") %>% 
  filter(NAME %in% c("Hennepin",
      "Dakota",
      "Carver",
      "Ramsey",
      "Anoka",
      "Scott",
      "Washington",
      "Wright",
      "Sherburne")) %>% 
    st_transform(4326)


segments <- st_intersection(segments, counti)

ggplot() +
  geom_sf(data = segments) +
  geom_sf(data = inrix_spatial,
          color = "cyan") +
  theme_minimal()
```

```{r}
inrix_segments <- st_intersection(segments, inrix_spatial)

ggplot() +
  geom_sf(data = inrix_spatial,
          color = "lightcyan4",
          size = 2) +
  geom_sf(data = inrix_segments,
          alpha = 0.5,
          color = "gray") +
  theme_minimal()
```


```{r}
segments_join <- st_join(inrix_spatial, inrix_segments)

ggplot() +
  geom_sf(data = segments_join) +
  theme_minimal()
```

